A Search for Circumstellar Companions Around Nearby White Dwarf Sirius b

### 2021-07-27 paper/paper.tex
update paper

\DeclareSIUnit\parsec{pc}
\DeclareSIUnit\au{AU}
\DeclareSIUnit\mj{M_J}
\DeclareSIUnit\ms{M_\odot}
\DeclareSIUnit\year{yr}

@@

@@

Nearby white dwarfs (within $\sim$\qty{100}{\parsec}) are compelling targets for direct imaging searches, despite the limited knowledge of planetary systems around evolved stars. \cite{burleigh_imaging_2002} suggests planets on initially wide (>\qty{5}{\au}) orbits around intermediate mass stars (\qtyrange{1}{8}{\ms}) will survive expansion during the red giant (RGB) phase and remain bound to the resultant white dwarf. During the asymptotic giant (AGB) phase, strong ionizing soft X-ray and UV emission will likely photoevaporate planetary atmospheres. In addition,  the mass loss of the star will adiabatically expand the orbit of the planet by a maximum factor of $M_{*,\mathrm{WD}}/M_{*,\mathrm{MS}}$ \citep{jeans_cosmogonic_1924}.

Theoretical calculations show that planets can survive the red giant and asymptotic giant stages \citep{burleigh_imaging_2002,jura_pollution_2008,nordhaus_orbits_2013}, and results from radial velocity and transit surveys suggest extrasolar planets are prevalent throughout the Milky Way \citep{cumming_keck_2008}.

However, only one planetary mass object has been detected around a white dwarf (WD) with imaging \citep[WD 0806-661;][]{luhman_discovery_2011}.

Directly imaging exoplanets is a powerful method in exoplanet science which explores a parameter space very complementary to the popular transit photometry and radial velocity methods. The red giant stage has cleared planets within a few \unit{au} greatly lowering the transit probability. White dwarfs have very few spectral lines to make high-precision radial velocity measurements from spectral lines. Imaging, though, benefits from the orbital expansion of the planet, and the cooling of the white dwarf greatly reduces the contrast between them \citep{burleigh_imaging_2002,gould_finding_2008}. The Degenerate Objects around Degenerate Objects (DODO) survey observed 29 white dwarfs with Gemini/NIRI and VLT/ISAAC and provided upper limits around $\sim$\qty{8}{\mj} beyond \qty{35}{\au} \citep{hogan_dodo_2009}.

In this work we study the closest white dwarf to the Sun, Sirius b, and provide upper mass limits based on our observations in the near-infrared. \autoref{sec:obs} describes our target and observing strategy, \autoref{sec:analysis} describes the processing and analysis techniques used, \autoref{sec:results} describes our results and we conclude with \autoref{sec:conclusion}.

Sirius is one of the closest star system to the Sun at \qty{2.7}{\parsec} away, consisting of an A1Vm star (a) and a WD with an age around \qty{225}{\mega\year} \citep{bond_sirius_2017,collaboration_gaia_2018}. There are no known planets around either star, and using constrained 3-body numerical simulations a potential companion around Sirius b would only be stable with a \qty{1.79}{\year} orbit, which corresponds to a circular orbit with \qty{1.5}{\au} separation. Sirius b has been studied in the \qty{10}{\micro\meter} regime recently by \cite{pathak_high_2021}, who achieve \qtyrange[]{1.8}{3.5}{\mj} upper limits within 2 arcseconds.

@@

We targeted Sirius B directly using Keck/NIRC2 in Lp-band across three epochs in 2020 (\autoref{tbl:obs}). Our first attempt to observe Sirius B failed due to the strong scattered light from Sirius A. The adaptive optics (AO) calibration routinely failed when the extremely bright (-1.35 K-band magnitude) diffraction patterns would sweep into the field of view (FoV) of Sirius B (shown in \autoref{fig:spike}). Similarly, trying to deploy the focal-plane vortex coronagraph \citep{serabyn_w_2017} failed when the low-order wavefront correction algorithm, QACITS \citep{huby_w_2017}, performed erratically in the presence of the scattered light. In order to overcome these issues, we decided to try using Sirius A as the NGS and off-axis guiding to Sirius B.

@@

Another metric we explored was the STIM largest intensity mask map \citep[SLIM map;][]{pairet_signal_2020}. The SLIM map is an ensemble statistic which calculates the average STIM map from many residual cubes along with the average mask of the $N$ most intense pixels in each STIM map. We use the SLIM map with the PCA, NMF, and GreeDS algorithms because they all depend on choosing the number of basis components, which is hard to determine \textit{a priori}. For each algorithm, we create residual cubes for increasing number of components, from 1 to 10. The collapsed residual frames, average STIM map, SLIM map, and contrast curves for each epoch for each of the above algorithms can be found in \autoref{sec:adi-results}. All of the code for these reductions can be found at the GitHub repository\footnote{\href{https://github.com/mileslucas/sirius-b}{https://github.com/mileslucas/sirius-b}}.

@@

The contrast maps from the best performing algorithm for each reduction are shown in \autoref{fig:contrast}. We determine the limiting sensitivities in terms of the planetary mass by first calculating the contrast-limited magnitude using an Lp-band magnitude for Sirius B of 9.1 (adapted from \citealp{bonnet-bidaud_adonis_2008}). Then we use an age of \qty{225}{\mega\year} to interpolate the planetary mass using the AMES-Cond evolutionary grid and atmosphere models \citep{allard_models_2012}. The best performing epoch was the 2020-11-21 night, which managed to reach an exceptional sensitivity of \qtylist[list-units=single]{3.5;1.9;1.1;0.72;0.63}{\mj} at projected separations of \qtylist[list-units=single]{0.25;0.5;0.75;1;2}{\au} (\qtylist[list-units=single]{0.09;0.19;0.28;0.38;0.75}{"}) respectively.

@@

It is interesting to note that there is a symmetric shape in the residual frames from GreeDS and NMF for the November epochs. Both these algorithms were designed for imaging disks, so it is natural to consider this structure to be some kind of circumstellar debris disk.

@@

\begin{figure*}
    \plotone{orbit}
    \caption{Provisional orbits fit using OFTI with $10^4$ orbits. None of the orbits contain all three points within their uncertainties.}
\end{figure*}

>>>>

\DeclareSIUnit\parsec{pc}
\DeclareSIUnit\au{AU}
\DeclareSIUnit\mj{M_J}
\DeclareSIUnit\ms{M_\odot}
\DeclareSIUnit\year{yr}
\sisetup{
    list-units=single
}

@@

We present the deepest study of planetary companions around Sirius B, the closest and brightest white dwarf in the night sky. We use Keck/NIRC2 in Lp band (\qty{3.776}{\micro\meter}) across 3 epochs in 2020 to take direct images of Sirius B. Each epoch was prepared and processed using the technique of angular differential imaging, which allowed us to reach exceptional sensitivity limits. The best-performing contrast from each epochs reached sensitivities of \qtylist{3.5;1.9;1.1;0.72;0.63}{\mj} at projected separations of \qtylist{0.25;0.5;0.75;1;2}{\au} (\qtylist{0.09;0.19;0.28;0.38;0.75}{\arcsecond}) respectively. These sensitivites are the best yet around Sirius B and are better than any limits reached around other evolved star. Despite the exceptional sensitivity of our observations, we do not report any additional companions around Sirius B.

@@

Nearby white dwarfs (within $\sim$\qty{100}{\parsec}) are compelling targets for direct imaging searches, despite the limited knowledge of planetary systems around evolved stars. \cite{burleigh_imaging_2002} suggests planets on initially wide ($>$\qty{5}{\au}) orbits around intermediate mass stars (\qtyrange{1}{8}{\ms}) will survive expansion during the red giant (RGB) phase and remain bound to the resultant white dwarf. Furthermore, \citet{nordhaus_orbits_2013} explain that close-in planets that escape the Roche limit of the expanding red giant can still be shredded by tidal forces, however the tidal forces on planets that readily escape engulfment are negligible. During the asymptotic giant (AGB) phase, strong ionizing soft X-ray and UV emission will likely photoevaporate planetary atmospheres. In addition, the mass loss of the star will adiabatically expand the orbit of the planet by a maximum factor of $M_{*,\mathrm{MS}}/M_{*,\mathrm{WD}}$ \citep{jeans_cosmogonic_1924}.

If the planet has managed to survive through the RGB and AGB phases, stars below \qty{8}{\ms} will enter the white dwarf cooling sequence, which will lead to a reduction in luminosity by $10^3-10^4$. This means the contrast between a white dwarf and a companion will be 3-4 orders of magnitude lower than the equivalent main sequence system. The combination of expanded orbits and lower contrast are beneficial for imaging, due to the limited inner working angles and sensitivities of ground-based instrumentation. The transit photometry method is very insensitive to planets around white dwarfs because the probability of a transit decreases with increasing separation. Similarly, the radial velocity method is limited to very massive planets on larger separations in addition to the extremely broad spectral features of a white dwarf which greatly limit the precision of radial velocity measurements.

The first search for substellar companions around white dwarfs was conducted by \citet{probst_infrared_1983} by looking for infrared (IR) excess in their spectral energy distributions (SED) using broadband photometry. This excess could likely be attributed to circumstellar dust or gas discs, which are indicators of the planetary formation history. Of the $\sim$ 100 white dwarfs they studied, no companions were found. The same method was applied by \citet{zuckerman_excess_1987} who found excess IR emission around white dwarf G29-38. Eventually this excess was determined to be from a dust disc \citep{telesco_observations_1990}, with the current interpretation associating the dust with accretion onto the white dwarf, polluting the stellar atmosphere \citep{koester_metals_1997}.

The only planetary-mass object imaged around a single white dwarf (WD 0806-661B) was discovered using Spitzer with a mass of \qty{7}{\mj} on a very wide \qty{2500}{\au} orbit \citep{luhman_discovery_2011}. The ``Degenerate Objects around Degenerate Objects'' survey \citep[DODO;][]{hogan_dodo_2009} observed 29 white dwarfs with Gemini/NIRI and VLT/ISAAC, reaching an average upper limit around $\sim$\qty{8}{\mj} beyond \qty{35}{\au}. The young white dwarf GD 50 was observed using the extreme AO instrument SPHERE at the VLT \citep{xu_extreme-ao_2015}, reaching sensitivity limits of \qty{4}{\mj} at \qty{6.2}{\au}. Sirius B has also been imaged, which will be discussed later. To date, there has been no direct images of an exoplanet around a white dwarf.

The Sirius system is the 7th closest to the sun at \qty{2.7}{\parsec}, consisting of Sirius A, an A1Vm star known for being the brightest in the night sky, and Sirius B, a DA2 white dwarf with a \qty{50}{\year} orbit \citep{bond_sirius_2017,collaboration_gaia_2018}. The proximity and brightness of the Sirius system make it one of the most well-studied systems throughout history \citep[see][]{wesemael_sirius_1982}.

The orbit of Sirius B was first studied by \citet{bessel_variations_1844}, who recognized wobbles in the proper motion of Sirius A caused by a ``dark satellite''. \citet{bond_companion_1862} reported and confirmed visual evidence of Sirius B by Alvan G. Clark and his father. \citet{adams_spectrum_1915} took the first spectral measurements of Sirius B and found it to be very similar to a main-sequence early A-type star despite the faintness, which we know know to be typical of white dwarf specra.

Initial astrometric perturbations suggested a 50 year orbital period \citep{auwers_orbit_1864}, with the first orbital solution including dynamic masses published by \citet{van_den_bos_orbit_1960}. \citet{gatewood_study_1978} improve upone the van den Bos orbital solution with over 300 more images, finding nearly the same dynamical masses for the binary. \citet{bond_sirius_2017} greatly refine the orbit using a compilation of historical data and HST data, which gives dynamical masses of \qty{2.063+-0.023}{\ms} and \qty{1.018+-0.011}{\ms} for A and B, respectively.

The precise dynamical masses are key for finding the age of Sirius; \citet{bond_sirius_2017} report a cooling age for Sirius B of \qty{126}{\mega\year} using isochrones. Studies of the initial-final mass relation (IFMR) of white dwarfs \citep{cummings_two_2016} lead to estimates of the progenitor mass of Sirius B to be between \qtyrange{5}{5.6}{\ms}, which when combined with stellar evolution codes yield total ages between \qtyrange{226}{228}{\mega\year} with an uncertainty of about $\pm$\qty{10}{\mega\year} \citep{bond_sirius_2017}.

A companion around Sirius B would be affected by the orbit of Sirius A, and this constrained three-body system has been studied numerically \citep{holman_long-term_1999}. \citet{bond_sirius_2017} calculate the longest period stable companion around Sirius B is \qty{1.79}{\year}, which corresponds to a \qty{1.5}{\au} circular orbit.

Some of the first modern studies searching for companions around Sirius B were \citet{schroeder_search_2000} and \citet{kuchner_search_2000}, who used the HST planetary camera and NICMOS cameras, respectively, looking for signal around \qty{1}{\micro\meter}, which combined had a sensitivity down to $\sim$\qty{10}{\mj} at \ang{;;2} (\qty{5.3}{\au}). \citet{bonnet-bidaud_adonis_2008} use the ground-based ESO/ADONIS instrument in JHKs to reach a sensitivty of $\sim$\qty{30}{\mj} at \ang{;;3} (\qty{7.9}{\au}). \citet{skemer_sirius_2011} used mid-IR (out to \qty{10}{\micro\meter}) observations from Gemini/T-ReCs which ruled out evidence for any infrared excess around Sirius B. \citet{thalmann_piercing_2011} use Subaru/IRCS at \qty{4.05}{\micro\meter} reaching detection sensitivities of \qtyrange{6}{12}{\mj} at \ang{;;1}. Very recently, \citet{pathak_high_2021} took simultaneous mid-IR observations (\qty{10}{\micro\meter}) at VLT/VISIR of Sirius A (through a coronagraph) and B. Because of the simultaneous observation, their contrast is split depending on which half of the FoV is being considered. Their average sensitivity is $\sim$\qty{2.5}{\mj} at $\sim$\qty{1}{\au}, and their best sensitivity (from the ``inner'' region) is $\sim$\qty{1.5}{\mj} at \qty{1}{\au}.

In this work we report our study of Sirius B using direct images from Keck/NIRC2. \autoref{sec:obs} describes our target and observing strategy, \autoref{sec:analysis} describes the processing and analysis techniques used, \autoref{sec:results} describes our results and we conclude with \autoref{sec:conclusion}.

@@

Despite Sirius B being the brightest white dwarf in the sky, it is still 10 magnitudes dimmer than Sirius A, which makes it it a very technically challenging target, especially on ground-based telescopes which have to deal with extra scaterring and diffraction from the atmosphere.

We targeted Sirius B directly using Keck/NIRC2 in Lp-band (\qty{3.776}{\micro\meter}) across three epochs in 2020 (\autoref{tbl:obs}). Our first attempt to observe Sirius B failed due to the strong scattered light from Sirius A. The adaptive optics (AO) calibration routinely failed when the extremely bright (-1.35 K-band magnitude) diffraction patterns would sweep into the field of view (FoV) of Sirius B (shown in \autoref{fig:spike}). Similarly, trying to deploy the focal-plane vortex coronagraph \citep{serabyn_w_2017} failed when the low-order wavefront correction algorithm, QACITS \citep{huby_w_2017}, performed erratically in the presence of the scattered light. This is very similar to the issues reported by \citet[\S2]{vigan_high-contrast_2015}  in their attempts to image Sirius B. In order to overcome these issues, we decided to try using Sirius A as the NGS and off-axis guiding to Sirius B.

@@

Another metric used was the STIM largest intensity mask map \citep[SLIM map;][]{pairet_signal_2020}. The SLIM map is an ensemble statistic which calculates the average STIM map from many residual cubes along with the average mask of the $N$ most intense pixels in each STIM map. We use the SLIM map with the PCA, NMF, and GreeDS algorithms because they all depend on choosing the number of basis components, which is hard to determine \textit{a priori}. For each algorithm, we create residual cubes for increasing number of components, from 1 to 10. The collapsed residual frames, average STIM map, SLIM map, and contrast curves for each epoch for each of the above algorithms can be found in \autoref{sec:adi-results}. All of the code for these reductions can be found at the GitHub repository\footnote{\href{https://github.com/mileslucas/sirius-b}{https://github.com/mileslucas/sirius-b}}.

@@

The contrast maps from the best performing algorithm for each reduction are shown in \autoref{fig:contrast}. We determine the limiting sensitivities in terms of the planetary mass by first calculating the contrast-limited magnitude using an Lp-band magnitude for Sirius B of 9.1 (adapted from \citealp{bonnet-bidaud_adonis_2008}). Then we use an age of \qty{226}{\mega\year} to interpolate the planetary mass using the AMES-Cond evolutionary grid and atmosphere models \citep{allard_models_2012}. It is important to note how the precision of the system age (see \autoref{sec:intro}) leads to quite accurate interpolation on the evolutionary grids. The best performing epoch was on the night of 2020-11-21, which managed to reach an exceptional sensitivity of \qtylist[list-units=single]{3.5;1.9;1.1;0.72;0.63}{\mj} at projected separations of \qtylist[list-units=single]{0.25;0.5;0.75;1;2}{\au} (\qtylist[list-units=single]{0.09;0.19;0.28;0.38;0.75}{\arcsecond}) respectively.

@@

In this work we demonstrate the results of high-contrast images of Sirius B in the near-IR. Our sensitivy limits are the best that have been reached for Sirius B, reaching \qty{3.5}{\mj} at \qty{0.25}{\au}. In the mid-IR previous works have only probed as close-in as $\sim$\qty{1}{\au}, where we reach a sensitivity of \qty{0.72}{\mj}. Despite the exceptional sensitivity of this study, we found no appreciable evidence for a companion around Sirius B, consistent with previous results.

We have published alongside this work the entire codebase used for pre-processing the data, doing the ADI reductions, and for generating every figure in this manuscript. We have also published our reduced datasets under an open license. We hope that this improves the reproducibility of the work as well as providing data for exploring new and different ADI algorithms.

This study also shows the capability for near-IR non-coronagraphic imaging for faint white dwarfs. Our sensitivity limits are deeper and closer than any other WD imaging results, even those using coronagraphs, thanks in no small part to the proximity of Sirius B.

It is interesting to note the morphology of the innermost $\sim$\ang{;;0.4} in the frames produced by GreeDS and NMF. Both of these algorithms have been shown to outperform traditional median and PCA subtraction for disc imaging. In the frames from each epoch, but particularly in the two November epochs, a symmetric ``barbell'' shape can be seen. Due to the nature of high-contrast imaging, it is extremely difficult to differentiate systematic noise from real signal in the speckle-limited regime, in addition there is no evidence for IR excess from a circumstellar disc. Follow-up work in the visible (e.g., Subaru/VAMPIRES, VLT/SPHERE) may be able to image such a disc.

As this study and others have probed deeper and closer to Sirius B, we must consider, too, the astrophysics of post-main-sequence evolution. As mentioned in \autoref{sec:intro}, the expansion of the star during the RGB phase places a limit on the innermost orbital separation that would escape engulfment or strong tidal forces. A study into the evolution of the Sirius B progenitor through the RGB phase could provide these limits; it is possible we have reached those limits observationally and could provide some definitive answer to whether Sirius B has a companion.

@@

\begin{figure*}
    \plotone{orbit_frames}
    \caption{Provisional astrometry (white circles) displayed on STIM maps using the GreeDS algorithm with 2 components. Each frame was been cropped to the inner \ang{;;0.25}. The inner FWHM has been masked out in each frame.}
\end{figure*}

\begin{figure*}
    \plotone{orbit_corner}
    \caption{Corner plot of the marginal posteriors from the OFTI samples.}
\end{figure*}
